Goal Recognition as Planning with Unknown Domain Models

### 2019-02-12 doc/ijcai-19-goalRecognition/goalRecognition.tex
processing Miquels comments for pages 1 and 2 ijcai19 paper on goal recognition

\subsection{The bias of the initially {\em empty} action model}
All $\{pre\_e\_\xi, eff\_e\_\xi\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}$ fluents are false at the initial state of our compilation. Classical planners usually preffer shorter solution plans, so a bias may be introduced to the $P=\tup{F,A[\cdot],I,G}$ task preferring solutions that imply action models with a shorter number of {\em preconditions}/{\em effects} (i.e., classical plans that solve $P'=\tup{F',A',I,G}$ using a shorter number of {\em insert} actions).

This bias could be eliminated defining a cost landscape for the actions in $A'$ (e.g. {\em insert} actions has {\em zero cost} while $\mathsf{apply_{\xi,\omega}}$ actions has a {\em positive constant cost}). In practice, since classical planners are not proficiency optimizing cost landscapes of this kind, we use a different approach to disregard the cost of the {\em insert} actions. Our approach is to use a SAT-based planner~\cite{rintanen2014madagascar} because it can apply all actions for inserting preconditions in a single planning step (these actions do not interact). Further, the actions for inserting action effects are also applied in a single planning step so the plan horizon for programming any action model is always bound to 2, which significantly reduces the planning horizon.

Our compilation for {\em planning with unknown domain models} can then be understood as an extension of the SATPLAN approach for classical planning~\cite{kautz1992planning} with two additional initial layers: a first one for inserting the action preconditions and a second one for inserting the action effects. These two extra layers are followed by the typical $N$ layers of the SATPLAN encoding (extended however to apply the action models that are determined by the previous two initial layers). Regarding again to the example of Figure~\ref{fig:plan-lplan}, this means that steps [00-04] are applied in paralel in the first SATPLAN layer, steps [05-13] are applied in paralel in the second layer and each step [14-17] is applied sequentially and correponds to a differerent SATPLAN layer (so the example would be solved using just six layers).

The SAT-based planning approach is also convenient for the task of {\em goal recognition as planning with unknown domain models} because its ability to deal with planning problems populated with dead-ends and because symmetries in the insertion of preconditions/effects into an action model do not affect to the planning performance.

@@

\begin{proof}[Proof]
\begin{small}
Once a given precondition or effect is inserted into an action model it can never be removed back and once an action model is applied it cannot be {\em reprogramed}. In the compiled problem the value of $F$, the fluents of the original problem, can exclusively be modified via $\mathsf{apply_{\xi,\omega}}$ actions.  The set of goals $G$ can only be achieved executing an applicable sequence of $\mathsf{apply_{\xi,\omega}}$ actions that, starting in the corresponding initial state reach a state $G \subseteq s_n$. This means that the action model used by the $\mathsf{apply_{\xi,\omega}}$ actions has to be consistent with the traversed intermediate states. We know that this must be true by the definition of the $\mathsf{apply_{\xi,\omega}}$ so hence, the sub-sequence of $\mathsf{apply_{\xi,\omega}}$ appearing in $\pi'$ to solve $P'$ is a solution plan to $P=\tup{F,A[\cdot],I,G}$. 
\end{small}
\end{proof}

@@

The size of the classical planning task $P'$ output by our compilation depends on the arity of the given {\em predicates} $\Psi$, that shape the propositional state variables $F$, and the number of parameters of the action models, $|pars(\xi)|$. The larger these arities, the larger $|{\mathcal I}_{\Psi,\xi}|$. The size of the ${\mathcal I}_{\Psi,\xi}$ set is the term that dominates the compilation size because it defines the $\{pre\_e\_\xi, eff\_e\_\xi\}$ fluents, the corresponding set of {\em insert} actions, and the number of conditional effects in the $\mathsf{apply_{\xi,\omega}}$ actions.

>>>>

\subsection{The bias of the initially {\em empty} action model}
Classical planners tend to preffer shorter solution plans, so our compilation may introduce a bias to $P=\tup{F,A[\cdot],I,G}$ problems preferring solutions that are referred to action models with a shorter number of {\em preconditions}/{\em effects}. In more detail, all $\{pre\_e\_\xi, eff\_e\_\xi\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}$ fluents are false at the initial state of our $P'=\tup{F',A',I,G}$ compilation so classical planners tend to solve $P'$ with plans that require a shorter number of {\em insert} actions.

This bias could be eliminated defining a cost function for the actions in $A'$ (e.g. {\em insert} actions has {\em zero cost} while $\mathsf{apply_{\xi,\omega}}$ actions has a {\em positive constant cost}). In practice we use a different approach to disregard the cost of {\em insert} actions because classical planners are not proficiency optimizing {\em plan cost} with zero-cost actions. Instead, our approach is to use a SAT-based planner~\cite{rintanen2014madagascar} because it can apply all actions for inserting preconditions in a single planning step (these actions do not interact). Further, the actions for inserting action effects are also applied in a single planning step so the plan horizon for programming any action model is always bound to 2, which significantly reduces the planning horizon.

Our compilation for {\em planning with unknown domain models} can then be understood as an extension of the SATPLAN approach for classical planning~\cite{kautz1992planning} with two additional initial layers: a first layer for inserting the action preconditions and a second one for inserting the action effects. These two extra layers are followed by the typical $N$ layers of the SATPLAN encoding (extended however to apply the action models that are determined by the previous two initial layers, the $\mathsf{apply_{\xi,\omega}}$ actions). Regarding again the example of Figure~\ref{fig:plan-lplan}, this means that steps [00-04] are applied in paralel in the first SATPLAN layer, steps [05-13] are applied in paralel in the second layer and each step [14-17] is applied sequentially and correponds to a differerent SATPLAN layer (so just six layers are necesary to compute the example plan of Figure~\ref{fig:plan-lplan}).

The SAT-based planning approach is also convenient for the task of {\em goal recognition as planning with unknown domain models} because its ability to deal with classical planning problems populated with dead-ends and because symmetries in the insertion of preconditions/effects into an action model do not affect to the planning performance.

@@

\begin{proof}[Proof]
\begin{small}
Once a given precondition (or effect) is inserted into an action model it can never be removed back and once an action model is applied (via an $\mathsf{apply_{\xi,\omega}}$ action) it cannot longer be {\em programed}. The set of goals $G$ can only be achieved executing an applicable sequence of $\mathsf{apply_{\xi,\omega}}$ actions that, starting in the corresponding initial state reach a state $G \subseteq s_n$ (the fluents of the original problem $F$ can exclusively be modified by the $\mathsf{apply_{\xi,\omega}}$ actions). This means that the action model used by the $\mathsf{apply_{\xi,\omega}}$ actions has to be consistent with the traversed intermediate states. We know that this must be true by the definition of the $\mathsf{apply_{\xi,\omega}}$ so hence, the sub-sequence of $\mathsf{apply_{\xi,\omega}}$ appearing in $\pi'$ to solve $P'$ is a solution plan to $P=\tup{F,A[\cdot],I,G}$. 
\end{small}
\end{proof}

@@

The size of the classical planning task $P'$ output by our compilation depends on the arity of the given {\em predicates} $\Psi$ and the number of parameters of the action models, $|pars(\xi)|$. The larger these arities, the larger $|{\mathcal I}_{\Psi,\xi}|$. This term dominates the compilation size because it defines the $\{pre\_e\_\xi, eff\_e\_\xi\}$ fluents, the corresponding set of {\em insert} actions, and the number of conditional effects of the $\mathsf{apply_{\xi,\omega}}$ actions.
